\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{float}
\graphicspath{{images/}}
\hypersetup{
  colorlinks=true,
  urlcolor=blue,
}
\begin{document}

\title{A Floating-Point Standard for Balanced Ternary}
\author{Aresh Pourkavoos}
\maketitle

\newcommand{\T}{\mathrm{T}}

Balanced ternary is a base 3 number system,
i.e. numbers in balanced ternary are written as a string of digits
whose contributions to the number are scaled by powers of 3.
However, rather than the usual 0, 1, and 2,
the digits of balanced ternary are $-1$, 0, and 1.
$-1$ is usually written as T, as it looks like a 1 with a bar or negative sign on top.
It can represent negative integers without the use of a negative sign,
e.g. $\T1 = (-1) \times 3 + 1 \times 1= -2$.

In the early days of computers,
this system was considered as an alternative to binary,
and many balanced ternary computers were built,
most notably the Setun in Russia in the early 1960s.
A brief side-by-side comparison follows.
First of all, since there is no real distinction between positive and negative integers,
there is no need for additional circuits to handle each case.
However, the same is also true for all modern binary computers
thanks to the two's complement implementation of signed integers,
which takes advantage of the fact that logic gates on a finite number of bits
behaves like modular arithmetic.
In two's complement, the larger half of what would be unsigned integers,
e.g. 8 through 15 (inclusive) for 4-bit integers,
are reassigned to negative numbers (-8 through -1 in this case).
So in this case, the advantage evens out.

Next, addition and multiplication in balanced ternary
both rely on their respective tables for a single digit.
In binary, adding two bits requires a carryover $\frac{1}{4}$ of the time,
namely when performing $1+1$.
In balanced ternary, however, this occurrs only $\frac{2}{9}$ of the time,
where $1+1=1T$ and $T+T=T1$,
giving it a marginal advantage. 

The efficiency of a base in representing numbers
can be calculated using radix economy,
which measures the limiting behavior
of the length of large numbers stored in base $b$,
where each digit is $b$ units long,
as if it were stored by a one-hot vector of bits.
In base $b$, the radix economy is given by $\frac{b}{\ln(b)}$,
and a lower radix economy is more efficient.
Binary and quaternary (base 4) both have
a radix economy of approximately $2.89$,
whereas base 3's is $2.73$ -
again, a slim advantage over binary.
However, the practical usefulness of radix economy is debated,
since it makes assumptions about the distribution
of numbers that computers typically store
which may not be in accordance with reality.

Where I believe balanced ternary shines is in how it handles floating point numbers.


Float $f$ comprised of exponent $e$ (int) and significand $s$ (-1.5 to 1.5) \\
If $s$ starts with 1 or T, $f = 3^e \times s$ (normal) \\
Else if $e=e_{min}$ (all Ts), $f$ is denormal (see formula for normal) \\
Else, $f =$ sNaN, qNaN, or $\pm\infty$ (TBD when operations are implemented) \\
0 is unique, no $\pm 0$ like in binary floats
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.25}

\noindent
Example with 1 trit for $e$ and 2 trits for $s$: \\
\hspace{-2cm}
\begin{tabular}{*{10}{|c}|}
  \hline
  Memory
  & $\T\T\T$ & $\T\T0$ & $\T\T1$
  & $\T0\T$ & $\T00$ & $\T01$
  & $\T1\T$ & $\T10$ & $\T11$ \\ \hline
  Formula
  & $10^\T \times \T.\T$ & $10^\T \times \T.0$ & $10^\T \times \T.1$
  & $10^\T \times 0.\T$ & $10^\T \times 0.0$ & $10^\T \times 0.1$
  & $10^\T \times 1.\T$ & $10^\T \times 1.0$ & $10^\T \times 1.1$ \\ \hline
  Result
  & $.\T\T$ & $.\T0$ & $.\T1$
  & $.0\T$ & $.00$ & $.01$
  & $.1\T$ & $.10$ & $.11$ \\ \hline
  Decimal
  & $-4/9$ & $-1/3$ & $-2/9$
  & $-1/9$ & $0$ & $1/9$
  & $2/9$ & $1/3$ & $4/9$ \\ \hline
  \hline
  Memory
  & $0\T\T$ & $0\T0$ & $0\T1$
  & $00\T$ & $000$ & $001$
  & $01\T$ & $010$ & $011$ \\ \hline
  Formula
  & $10^0 \times \T.\T$ & $10^0 \times \T.0$ & $10^0 \times \T.1$
  & ? & ? & ?
  & $10^0 \times 1.\T$ & $10^0 \times 1.0$ & $10^0 \times 1.1$ \\ \hline
  Result
  & $\T.\T$ & $\T.0$ & $\T.1$
  & ? & ? & ?
  & $1.\T$ & $1.0$ & $1.1$ \\ \hline
  Decimal
  & $-4/3$ & $-1$ & $-2/3$
  & ? & ? & ?
  & $2/3$ & $1$ & $4/3$ \\ \hline
  \hline
  Memory
  & $1\T\T$ & $1\T0$ & $1\T1$
  & $10\T$ & $100$ & $101$
  & $11\T$ & $110$ & $111$ \\ \hline
  Formula
  & $10^1 \times \T.\T$ & $10^1 \times \T.0$ & $10^1 \times \T.1$
  & ? & ? & ?
  & $10^1 \times 1.\T$ & $10^1 \times 1.0$ & $10^1 \times 1.1$ \\ \hline
  Result
  & $\T\T.$ & $\T0.$ & $\T1.$
  & ? & ? & ?
  & $1\T.$ & $10.$ & $11.$ \\ \hline
  Decimal
  & $-4$ & $-3$ & $-2$
  & ? & ? & ?
  & $2$ & $3$ & $4$ \\ \hline
\end{tabular}

Note that in order to avoid a special case for denormal numbers,
this format effectively wastes a third of possible code words.
It is possible to do the same for binary floats,
storing the extra 1 in the mantissa which is usually omitted
and changing it to 0 for denormal numbers,
but the IEEE decided against it.

\end{document}
